---
title: "Modeling Demographic Stochasticity and Thermal Mismatch in Host-Pathogen Wildlife Systems"
author: "Andy Carlino"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup:

Working Directory and Libraries:
```{r}
# Working Directory
setwd("C:\\Users\\andre\\OneDrive\\Documents\\GitHub\\DemoStoch_ThermMismatch_EpiModels")

# Libaries
library(tidyverse)
library(tidyr)
library(ggplot2)
library(stats)
library(reshape2)
library(gridExtra)
```

Source Functions:
```{r}
# Tau leap function
source("tau_leap_function.R")

# Simulation of the within-host model function
source("model_simulation_function.R")
```


# Quadratic Functions for Parameter Values along Thermal Gradient:

Thermal Performance Curve using a Quadratic function: 
param = (-q * [temperature - t_min][temperature - t_max])

```{r}
# Generate temperature range
temperatures <- seq(2, 20, length.out = 50)
```

Estimating the Range of Parameter Phi (Viral Replication Rate)
```{r}
estimate_phi <- function(temperature, q_phi, tmin_phi, tmax_phi) {
  # Quadratic equation
  phi <- pmax(-q_phi * (temperature - tmin_phi) * (temperature - tmax_phi), 0) 
  return(phi)
}

# Set q for phi
q_phi <- 0.01275
# Set temperature range
tmin_phi <- 1
tmax_phi <- 28.4

# Calculate corresponding values of phi
phi_values <- estimate_phi(temperatures, q_phi, tmin_phi, tmax_phi)

# Peak value of the phi Parameter
temp_peak_phi <- (tmin_phi + tmax_phi) / 2
peak_phi <- estimate_phi(temp_peak_phi, q_phi, tmin_phi, tmax_phi)
```


Estimating the Range of Parameter Alpha (Mass-action Attack Rate)
```{r}
estimate_alpha <- function(temperature, q_alpha, tmin_alpha, tmax_alpha){
  # Quadratic Equation
  alpha <- pmax(-q_alpha * (temperature - tmin_alpha) * (temperature - tmax_alpha), 0)
  return(alpha)
}

# Set q value for alpha 
q_alpha <- 0.00855
# Set temperature range
tmin_alpha <- -5
tmax_alpha <- 23.6

# Calculate corresponding values of alpha
alpha_values <- estimate_alpha(temperatures, q_alpha, tmin_alpha, tmax_alpha)

# Peak value of the alpha Parameter
temp_peak_alpha <- (tmin_alpha + tmax_alpha) / 2
peak_alpha <- estimate_alpha(temp_peak_alpha, q_alpha, tmin_alpha, tmax_alpha)
```


Estimating the Range of Parameter Psi (Immune Growth Rate in Response to Virus)
```{r}
estimate_psi <- function(temperature, q_psi, tmin_psi, tmax_psi){
  # Quadratic Equation
  psi <- pmax(-q_psi * (temperature - tmin_psi) * (temperature - tmax_psi), 0)
  return(psi)
}

# Set q value for psi 
q_psi <- 0.0041
# Set temperature range
tmin_psi <- 1
tmax_psi <- 28

# Calculate corresponding values of alpha
psi_values <- estimate_psi(temperatures, q_psi, tmin_psi, tmax_psi)

# Peak value of the alpha Parameter
temp_peak_psi <- (tmin_psi + tmax_psi) / 2
peak_psi <- estimate_psi(temp_peak_psi, q_psi, tmin_psi, tmax_psi)
```

Plot the functions: 
```{r}
plot_data <- data.frame(Temperature = temperatures,
                        Phi = phi_values, 
                        Alpha = alpha_values)
                        #Psi = psi_values)

# Plot Phi Values
phi_plot <- ggplot(plot_data, aes(x = Temperature, y = Phi)) +
  geom_line(color = "red") + 
  labs(x = "Temperature (C)", y = "Phi (Pathogen) Values") + 
  geom_vline(xintercept = temp_peak_phi, color = "red", linetype = "dashed") + 
  theme_bw() +
  ggtitle("Temperature vs. Parameter Values") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(size = 20)) + 
  theme(axis.text = element_text(size = 15)) + 
  theme(axis.title = element_text(size = 15))

# Plot Alpha Values
alpha_plot <- ggplot(plot_data, aes(x = Temperature, y = Alpha)) +
  geom_line(color = "blue") + 
  labs(x = "Temperature (C)", y = "Alpha (Host) Values") + 
  geom_vline(xintercept = temp_peak_alpha, color = "blue", linetype = "dashed") +
  theme_bw() + 
  theme(axis.text = element_text(size = 15)) + 
  theme(axis.title = element_text(size = 15))

# Plot Psi Values
#psi_plot <- ggplot(plot_data, aes(x = Temperature, y = Psi)) + 
#  geom_line(color = "green") + 
#  labs(x = "Temperature (C)", y = "Psi (Immune Response to Virus) Values") + 
#  geom_vline(xintercept = temp_peak_psi, color = "green", linetype = "dashed") +
#  theme_bw() + 
#  theme(axis.text = element_text(size = 15)) + 
#  theme(axis.title = element_text(size = 15))

plot_data_long <- plot_data %>%
  pivot_longer(cols = c(Phi, Alpha), 
               names_to = "Parameter", 
               values_to = "Value")

# Create the combined plot with a legend
combined_plot <- ggplot(plot_data_long, aes(x = Temperature, y = Value, color = Parameter)) +
  geom_line() +
  geom_vline(data = data.frame(Parameter = c("Phi", "Alpha"),
                               Peak = c(temp_peak_phi, temp_peak_alpha)),
             aes(xintercept = Peak, color = Parameter), linetype = "dashed", linewidth = 1.2) +
  labs(x = "Temperature (C)", y = "Parameter Values", 
       title = "Temperature vs. Parameter Values") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 15),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12)) +
  scale_color_manual(values = c("Phi" = "red", "Alpha" = "blue"), 
                     labels = c("Phi (Pathogen)", "Alpha (Host)")) +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.position = "right")

combined_plot
```

